Developing a preprint culture in biology

### 2012-10-05 ms.tex
merge Dom, improve Git and figshare sections

We briefly discuss the main options to submit preprints to open servers:
arXiv.org, Figshare, and the upcoming PeerJ and F1000Research.

@@

\subsection{Figshare}

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify each item. All content can be
commented and is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0. The CC-BY license grants rights to
\emph{copy, distribute, display and perform the work and make derivative works
based on it only if they give the author or licensor the credits in the manner
specified by these.}  CC0 is the most permissive license and effectively puts
the work in public domain (no rights reserved) or, if it is not possible in the
given jurisdiction, provides a simple permissive license.

@@

PeerJ uses the CC-BY-SA 3.0 license, which is similar to the CC-BY license used
by Figshare but adds the \emph{Share-alike} (sa) restriction that derivate works
need a license identical to the license that governs the original work.

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, Figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually
be reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee % What about corrections? There are exchanges between referees and the authors?
response. F1000Research works closely with data providers to integrate raw
data to the paper. For instance, upon submitting a paper, authors are asked to
upload their data, which are then integrated in \emph{e.g.} FigShare widgets,
the DOI of which are given in the paper when the data are first mentionned.
The licensing of the data is similar to the one used by FigShare, meaning that
the articles are free to access, and can be redistributed readily. By putting
much effort in integrating data to the paper, F1000Research is working to make
science more reproducible and open.

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous contributers to work asynchronously on the same project, often in parallel branches, all of which can be effortlessly merged and version controlled. Although created primarily for software development, git is ideal for academic research since it provides a way to version control and collaborate on every step of the manuscript development process, from data manipulation and analysis to writing and revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then merge their changes into a master branch.This takes the preprint process to an entirely new level, where the entire writing process is open from the beginning.
% I would avoid using the term forking since it is not a central piece of the section. It is one specific way of collaborating on a github repo when everyone does not have rights to it. People could just as easily work on a single branch, all with write access, either directly to the master branch or to individual branches that could be merged. So in other words, we should keep the language general so people don't get bogged down with trying to understand what a fork is.
% also I changed one author clones, then submits pull request langauge a bit since the main point of using git is that we can all work simultaneously unlike svn.
% I took out "These techniques are widely used in open source software development." because git can also be used for closed source software, offering the same advantages, where the repositories are private. Paid git accounts allow for private repos.

@@

% I added the following three paragraphs to the conclusion, offering also a discussion on the on going explosion of different publication models. Take what you want and remove what you don't need.
% I also had the feeling we might want to include a single paragraph, very short, calling that evaluations (academia, grant agencies etc...) looking at publication records or report of scientific activities start to view preprints as a significant and sound scientific contribution. One of the most important benefit is again the speed of the publication process. Most grants are for three years or less. It makes it hard to have a published paper within the limits of the grant. Preprints could however speed the process and show that contributions were made during the time of the grant. 
The ongoing discussions on the publication process, peer-reviewing and alternative publication models are all symptoms of the current uneasiness of scientists with the ever growing obsession with bibliographic metrics such as the impact factor \cite{Fisher2012}. There is pressure on researchers, wether they are junior looking for an academic position or senior trying to secure funding for their team, to orient their publication strategy in order to maximize their number of publications and total citations. A well-known consequence is to submit manuscripts first to the most prestigious journals, and then resubmit to lower level journals as they are rejected. The numerous negative impacts of such behavior have been discussed in depth \cite{hoc09}. They all contribute somehow in a general slowing down of scientific progress. Research activities and the publication process are drifting away from their fundamental object, namely the diffusion of novel scientific discoveries.

@@

% Fisher, J, Ritchie, E.G., Hanspach, J. (2012). Academia's obsession with quantity. Trends in Ecology and Evolution, 27, 473-474.
% Aarssen, L. (2012). Are peer-review filters optimal for the progress of science in ecology and evolution? Ideas in Ecology and Evolution, 5, 9–12.

@@

% Aren't we bridging more than cultural and geographic? I mean, we are essentially cross pollinating ideas. e.g. It took a lot time for MCMC to become popular in ecology even though it was widely used in other disciplines a decade or two earlier.
Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. Examples include a recent series of
papers on the theory of natural selection that was posted to arXiv
simultaneously with its publication in the \emph{Journal of Evolutionary
Biology} \cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}. Other
submissions in this category include evolutionary and ecological theory by
authors trained in physics and computer science.  Since authors in these fields
regularly check arXiv, submitting preprints may be the most effective way
biologists can help others avoid ``repeated work'' \cite{de2011contribution} and
form a synthetic community of evolutionary theorists from disparate backgrounds.
The advantages are clear and the costs are low.

@@

We thank Carl Boettiger for helpful comments on an earlier version of this
manuscript.

>>>>

We briefly discuss the main options to submit preprints to open servers:
arXiv.org, figshare, and the upcoming PeerJ and F1000Research.

@@

\subsection{figshare}

figshare (\href{http://figshare.com}{http://figshare.com}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify each item. All content can be
commented and is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0. The CC-BY license grants rights to
\emph{copy, distribute, display and perform the work and make derivative works
based on it only if they give the author or licensor the credits in the manner
specified by these.}  CC0 is the most permissive license and effectively puts
the work in public domain (no rights reserved) or, if it is not possible in the
given jurisdiction, provides a simple permissive license.

@@

PeerJ uses the CC-BY-SA 3.0 license, which is similar to the CC-BY license used
by figshare but adds the \emph{Share-alike} (sa) restriction that derivate works
need a license identical to the license that governs the original work.

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee
response. F1000Research works closely with data providers to integrate raw data
to the paper. For instance, upon submitting a paper, authors are asked to upload
their data, which are then integrated in \emph{e.g.} figshare widgets, the DOI
of which are given in the paper when the data are first mentionned.  The
licensing of the data is similar to the one used by figshare, meaning that the
articles are free to access, and can be redistributed readily. By putting much
effort in integrating data to the paper, F1000Research is working to make
science more reproducible and open.

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous
contributers to work asynchronously on the same project, often in parallel
branches, all of which can be effortlessly merged and version controlled.  While
Git is created primarily for software development, where the use of version
control systems is standard \cite{aru12}, it is ideal for academic research
since it provides a way to collaborate on every step of the manuscript
development process, from data manipulation and analysis to writing and
revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then
merge their changes into a master branch. This takes the preprint process to an
entirely new level, where the entire writing process is open from the beginning.

@@

% I added the following three paragraphs to the conclusion, offering also a
% discussion on the on going explosion of different publication models. Take
% what you want and remove what you don't need.  I also had the feeling we might
% want to include a single paragraph, very short, calling that evaluations
% (academia, grant agencies etc...) looking at publication records or report of
% scientific activities start to view preprints as a significant and sound
% scientific contribution. One of the most important benefit is again the speed
% of the publication process. Most grants are for three years or less. It makes
% it hard to have a published paper within the limits of the grant. Preprints
% could however speed the process and show that contributions were made during
% the time of the grant. 
The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers, wether
they are junior looking for an academic position or senior trying to secure
funding for their team, to orient their publication strategy in order to
maximize their number of publications and total citations. A well-known
consequence is to submit manuscripts first to the most prestigious journals, and
then resubmit to lower level journals as they are rejected. The numerous
negative impacts of such behavior have been discussed in depth \cite{hoc09}.
They all contribute somehow in a general slowing down of scientific progress.
Research activities and the publication process are drifting away from their
fundamental object, namely the diffusion of novel scientific discoveries.

@@

@@

% Aren't we bridging more than cultural and geographic? I mean, we are
% essentially cross pollinating ideas. e.g. It took a lot time for MCMC to
% become popular in ecology even though it was widely used in other disciplines
% a decade or two earlier.
Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. Examples include a recent series of
papers on the theory of natural selection that was posted to arXiv
simultaneously with its publication in the \emph{Journal of Evolutionary
Biology} \cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}. Other
submissions in this category include evolutionary and ecological theory by
authors trained in physics and computer science.  Since authors in these fields
regularly check arXiv, submitting preprints may be the most effective way
biologists can help others avoid ``repeated work'' \cite{de2011contribution} and
form a synthetic community of evolutionary theorists from disparate backgrounds.
The advantages are clear and the costs are low.

@@

We thank Carl Boettiger and Mark Hahnel for helpful comments on an earlier
version of this manuscript.
